---
title: "05 MetaAnalysis"
author: "Roxana Andreea Moldovan Moldovan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_float: false
    number_sections: false 
    code_folding: hide
    theme: flatly
    bg: "#FFFFFF"
    fg: "#202020"
    primary: "#90C432"
    secondary: "#2494b5" 
  pDT_document:
    extra_dependencies: "subfig"
linkcolor: blue
urlcolor: blue
citecolor: blue
header-includes:
- \usepackage{subfig}
- \usepackage{float}
fig_caption: yes 
link-citations: TRUE 
---
        
##   {.tabset .tabset-fade -} 
&nbsp;  


```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
# Load packages 
library(pacman)
pacman::p_load(glue)
pacman::p_load(DT)
pacman::p_load(magick)
```  


This document shows the results of the meta-analysis of the studies: GSE20950, GSE141432 para el contraste T2D - Ob.  
&nbsp;  


Taking a p-value adjusted for BH < 0.05 we found 1981 significative genes.  
&nbsp;  


### Table of results {-} 
&nbsp;  


```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
DT = get(load("C:/Users/roxya/OneDrive/Documentos/01Master_bioinformatica/00TFM/05Paper/AnalysisGit/Data/MA/Diabetes/ID/Meta-analysis_ID_DF.RData"))
datatable(DT, options = list(caption = "Meta-analysis results.",
                              scrollX = TRUE),
                              filter = "top")
```  


&nbsp;  


### Graphical representation {-}  
&nbsp; 

Here are the characteristic figures of the meta-analysis for the 
      significant genes taking a p-value adjusted for BH < 0.05.  
&nbsp;  


```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
# Plot function
grid_img <- function(list_fig, dir, patt, dim = 300){
  # Abrimos todas las filas:
  forest_plots = c()
  funnel_plots = c()
  influ_plots = c()
  for (fig in list_fig){
    f1 = glue("{fig}forest.svg")
    out1 <- magick::image_read(paste0(dir, "/", f1))
    forest_plots = c(forest_plots, image_scale(out1, dim))
    
    f2 = glue("{fig}funnel.svg")
    out2 <- magick::image_read(paste0(dir,"/", f2))
    funnel_plots = c(funnel_plots, image_scale(out2, dim))  
    
    f3 = glue("{fig}influence.svg")
    out3 <- magick::image_read(paste0(dir,"/", f3))
    influ_plots = c(influ_plots, image_scale(out3, dim))
  } 
  
  lim = length(forest_plots)
  grid = NULL
  for ( i in seq(list_fig)){
    a = forest_plots[[i]]
    b = funnel_plots[[i]]
    c = influ_plots[[i]]
    
    fila <- image_append(c(a, b, c))
    
    if(is.null(grid)){
      grid = fila
    }else{
      grid = image_append(c(grid,fila),  stack = TRUE)
    } 
  }
    return(grid)
}
```  


&nbsp;  


```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE, out.width=".99\\linewidth"}
path = "C:/Users/roxya/OneDrive/Documentos/01Master_bioinformatica/00TFM/05Paper/AnalysisGit/Data/MA/Diabetes/ID/Plots"
files <- list.files(path = path, pattern = "Meta-analysis_ID.*forest")
figuras = sapply( files, function(x) substr(x, start = 0, stop = nchar(x)-10), 
                  USE.NAMES = FALSE)
# Plot
grid_img(figuras, dir = path)
```  


&nbsp;  

