---
title: "05MetaAnalysisReport"
author: "Roxana Andreea Moldovan Moldovan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_float: false
    number_sections: false 
    code_folding: hide
    theme: flatly
    bg: '#FFFFFF'
    fg: '#202020'
    primary: '#90C432'
    secondary: '#2494b5' 
  pdf_document:
    extra_dependencies: "subfig"
linkcolor: blue
urlcolor: blue
citecolor: blue
header-includes:
- \usepackage{subfig}
- \usepackage{float}
fig_caption: yes 
link-citations: TRUE

---  

# Meta-Analysis summary {.tabset .tabset-fade -}

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
# Packages
pacman::p_load(ComplexHeatmap)
pacman::p_load(ComplexUpset)
pacman::p_load(dplyr)
pacman::p_load(flextable)
pacman::p_load(ggforce)
pacman::p_load(ggplot2)
pacman::p_load(ggVennDiagram)
pacman::p_load(glue)
pacman::p_load(officer)
pacman::p_load(sf)
pacman::p_load(stringr) 


Niv = c("IO", "IOM","IOF","SDIO",
        "ID", "IDM","IDF","SDID")
p_lim = 0.05

# Function
source("./Functions/Functions.R")
# Output table
RX_table <- function(Data,
                     Color1 = "#007FA6",
                     Color2 = "#007FA6" ,
                     Color.l = "black" ,
                     Color.l2 = "#878787",
                     Color.up = "#CB326D",
                     Color.down = "#00AFBB",
                     footer = NULL,
                     font = "Calibri"){
  cols=colnames(Data)[-c(1)] # Columns with info
  col_keys = colnames(Data) # Unique names
  # Header
  header = data.frame(col_keys = col_keys,
                      H1 = c("", rep("Contrast", length(cols))),
                      H2 = c("", cols)
                      )
  
  ## Flextable
  # Set info
  ft <- flextable(Data, col_keys = colnames(Data))
  # Set header
  ft <- set_header_df(ft, mapping = header, key = "col_keys" )
  # Format number
  ft <- colformat_num(ft, big.mark = ".", decimal.mark = "," )
  # Set footer
  if(! is.null(footer)){
    ft <- add_footer(ft, values = footer) %>% 
      merge_h(part = "footer")  %>% 
      bold(j=1, part = "footer") %>%
      hline(border = fp_border(width = 2, color = Color1), part = "footer")
  }
  
  # Merge headers
  ft <- merge_h(ft, part = "header") %>% 
    merge_v( part = "header")
  
  # Merge Contrasts
  #ft <- merge_v(ft, j = c(1,2), part = "body")
  
  # Horizontal lines 
  ft <- hline(ft,border = fp_border(width = 2, color = Color1), part = "header")
  
  # Outer bottom
  ft <- hline_bottom(ft, border = fp_border(width = 2, color = Color1))
  
  # Alignment
  ft <- flextable::align(ft, i=1, align = "center", part = "header") %>%
    flextable::align(i=2, align = "right", part = "header")
  ft <- flextable::align(ft, align = "right", part = "footer")
  ft <- flextable::align(ft, j=1, align = "left", part = "footer")
  
  
  # Header design
  ft <- fontsize(ft, i = 1:2, size = 12, part = "header") %>%
    color(i=1, color = Color2, part = "header") %>%
    color(i=2, color = Color.l2, part = "header") %>%
    bold(i=1, bold = TRUE, part = "header")
  # Rotate names
  #ft <- rotate(ft, i = 2, rotation = "btlr", part = "header", align = "bottom")
  
  # Font
  #font(ft, fontname = font, part = "all")
  
  # Color matrix
  colormatrix <- ifelse(Data[, cols] == 0, Color.l2, Color.l )
  # Color
  ft <- color(ft, j=cols, color = colormatrix, part = "body") 
  # Color
  x = Data[, 1]
  colors = case_when(
    x == "Up" ~ Color.up,
    x == "Down" ~ Color.down,
    TRUE ~ Color.l
  )
  ft <- color(ft, j=1, color = colors, part = "body") %>%
    bold(j=1, bold = TRUE, part = "body")

  return(ft)
} 

RXcolor <- function(lis, max=50){ 
  color = "black"
    add = NULL
    n = length(lis)
    if(n > max){
      lis = lis[1:max]
      add = glue("(+{n - max})")
    }
  ret = sapply(lis[-length(lis)], function(x){
    if(x %in% ObD & x %in% DM2D){
      color = "#5aac73"
  }else if(x %in% DM2D){color = "#2494B5"
    }else if(x %in% ObD){color = "#90C432"
    }else{color = "black"}
  ret = glue("as_chunk('{x}', props = fp_text_default(color = '{color}')), as_chunk(', ', props = fp_text_default(color = 'black'))")
  return(ret)
})
  # Last one
  ret = c(ret, glue("as_chunk('{lis[length(lis)]}', props = fp_text_default(color = '{color}'))"))
  # Add rest
  if (!is.null(add)){ret = c(ret, glue("as_chunk(', ', props = fp_text_default(color = 'black')), as_chunk('{add}', props = fp_text_default(color = 'black'))"))}
  return(ret)
}


RX_venn <- function(lists,
                    colors,
                    names,
                    text_color = "#878787",
                    title = NULL){
  venn <- Venn(lists)
  d <- process_data(venn)
  d2 <- process_data(venn)
  
  d2@region <- st_polygonize(d@setEdge) 
  plot = ggplot() +
  geom_sf(aes(fill = name), data = venn_region(d2)) +
  geom_sf(#aes(color = name), 
    size = 4, color = text_color,
    data = venn_setedge(d))  +
  geom_sf_text(aes(label = count), data = venn_region(d),
               color = text_color, size = 4.5) +
  scale_fill_manual(values = alpha(colors, .3)) +
  theme_void() +  # Tema
        theme(plot.background = element_rect(fill="transparent", color="transparent"),
              panel.background = element_rect(fill = "transparent", color="transparent"),
              axis.line = element_line(colour = "transparent"),
              legend.position = "bottom"
              #legend.position = "rigth"
              ) +
    guides(fill = guide_legend(title = NULL,
                                order = 1)) +
  if(! is.null(names)){
      plot = plot +
        geom_sf_text(aes(label = name), data = venn_setlabel(d),
                     nudge_x = -0.1, nudge_y = -0.1) 
  }
    if(! is.null(title)){
      plot = plot + labs(title=title, ) + 
        theme(plot.title = element_text(hjust = 0.5))
    }
  return(plot)
  } 
```  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
Dir = "./Data/OpenTargets"
ObDB = read.table(glue("{Dir}/Ob.tsv"), sep = "\t", header = TRUE) 
ObD = ObDB$symbol
DM2DB = read.table(glue("{Dir}/DM2.tsv"), sep = "\t", header = TRUE)
DM2D = DM2DB$symbo

```  


## Obesity  {.tabset .tabset-fade -}  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
DirData = "./Data/MA/Obesity"

f= c("IO/Meta-analysis_IO_DF.RData", 
     "IOM/Meta-analysis_IOM_DF.RData",
     "IOF/Meta-analysis_IOF_DF.RData",
     "SDIO/Meta-analysis_SDIO_DF.RData")

files = glue("{DirData}/{f}")

```  


### Resumen {-}   

Summary of the meta-analysis in Obesity. 

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
# Prepare data
Datas = sapply(files, function(x) get(load(file=x)), simplify = FALSE)
p_lim = 0.05

Res = sapply(Datas, function(Data){
  Data <-  Data %>% 
    mutate(Significance = 
             case_when(
               logFC > 0 ~ "Up",
               logFC < 0 ~ "Down",
               TRUE ~ "-"
             ))
  res = table(Data[Data$p.adjust.BH < p_lim, "Significance"])
  return(c(res, Total = sum(res)))
}, simplify = FALSE)

contrasts2 = Niveles = Niv = c("IO", "IOM","IOF","SDIO")

names(Res) = contrasts2
```  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
Data = data.frame(Contrast = names(Res[[1]]),
                  Res, check.names = FALSE)
footer = c("Total number of genes", sapply(Datas, function(x) nrow(x)))
col_keys = colnames(Data)
names(footer) = col_keys
RX_table(Data, footer = footer) %>% width( width = 1.2, unit = "in")
Dataob = Data
```


### Intersection {.tabset .tabset-fade -}   

Study of the intersections between the genes obtained in the meta-analysis. 

#### Upset plot {-} 

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.align = "center" }
# Prepare data
p_lim = 0.05
logFC_lim = 0

# Gene lists
Direction = c("Up", "Down")

Data0 = sapply(Datas, function(Data){
  Up = rownames(Data[which(Data$p.adjust.BH < p_lim & Data$logFC > logFC_lim), ])
  Down = rownames(Data[which(Data$p.adjust.BH < p_lim & Data$logFC < 0), ])
  return(c("Up" = list(Up), "Down" = list(Down)))
}, simplify = FALSE)

names(Data0) = contrasts2 
```   


```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=10  }
MinSize = 0
An = unlist(Data0, recursive = FALSE) 
An = ComplexHeatmap::list_to_matrix(An)  

Nom = do.call(str_replace_all,
        c(list(colnames(An)), pattern = "\\.Up", replacement=" logFC>0"))
Nom = do.call(str_replace_all,
        c(list(Nom), pattern = "\\.Down", replacement=" logFC<0"))

Niv = c("ID, logFC>0", "ID, logFC<0", 
        "IDM, logFC>0", "IDM, logFC<0",
        "IDF, logFC>0", "IDF, logFC<0",
        "SDID, logFC>0", "SDID, logFC<0")
colnames(An) = Niv
Data = data.frame(An, check.names = FALSE)[,length(colnames(An)):1]

Direction = data.frame(set = colnames(Data),
                       #dir = rep(c("Up", "Down"), 4))
                       dir = rep(c("logFC > 0", "logFC < 0"), 4))

# Legend:
#Anots = as.data.frame(make_comb_mat(An)) 
#Leg = sapply(as.list(Anots), function(x) colnames(Data)[which(x == 1)], simplify = FALSE)
Legend = lapply(colnames(Data), function(x) upset_query(group= x,
                                                color="#878785",
                                                fill="#878787"))

#colnames(Data) = Names

UpsetPlot <- upset(Data, 
                   colnames(Data), 
                   name = "Intersecciones", guides = "over",
                   # Numero de interacciones minimas 
                   min_size =0,
    width_ratio=0.25,
    #stripes='transparent',
    # Sección horizontal
    stripes = upset_stripes(
      mapping=aes(color= dir),
      colors = c(
        "logFC > 0"= "grey89",#"#f2c4d6",
        "logFC < 0" = "white" #"#cbf0f2"
      ),
      data = Direction
    ), 
    matrix = intersection_matrix(
            geom=geom_point(
                shape=15,
                size=4,
                stroke = 1
            ), 
            segment=geom_segment(
              size= 1.2,
              arrow.fill = "#878787"
            ),
            outline_color=list(
                active='#878787',
                inactive='gray'
            )), #+ theme(text=element_text(size = 29,angle=0))
    
    # Colour of intersecting lines:
    # Assignment of colours
    #queries= Legend, 
    
    
    # We change the position of the set size
    set_sizes = upset_set_size(position='right',
                               #geom = geom_bar(color="red")
                               ) +    
      #geom_text(aes(label=..count..), #hjust=1.1,
      #         hjust=0, stat='count', color="white") + 
      ylab('Nº de genes significativos') + 
      theme(axis.text.x=element_text(size = 10,
                                     angle=0),
            text = element_text(size=10),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            axis.line.x = element_line(color="#878787"),
            axis.ticks = element_line(color="#878787"),
            panel.background = element_rect(fill = "white",
                                          color = "white")),
    # Interactions
    base_annotations = list(
        'Intersection size'=( 
            intersection_size(
              mapping=aes(fill="bars_color"),
              alpha = 0.2,
              color = "#90C432", # color barra
              width = 0.6, # Ancho barra
              text = list(
                vjust=-0.5,
                #vjust=0.5,
                #hjust=-0.2,
                #angle=90,
                size =4.5
            ), 
            text_colors=c(
                on_background='#878787', on_bar="#878787"
            )
            ) + scale_fill_manual(values=c("bars_color"="#90C432"), 
                                  guide="none") + # fill barra
              scale_alpha_manual(values = c(0.2))
            + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(),
               plot.background = element_rect(fill="transparent",
                                                   color = "transparent"),
          panel.background = element_rect(fill = "transparent",
                                          color = "transparent"),
          axis.line.y = element_line(color="#878787"),
          axis.text.y=element_text(size = 10, angle=0)
          #axis.line.x = element_line(color="black"),
          #axis.ticks = element_line(color="black")
          )
            + ylab('Genes significativos por grupo')
        )),
    sort_sets = FALSE
        
)  

UpsetPlot

```

#### Venn plot {-} 

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=4, fig.align='center'}
# Up
colors = c( "ID" = "#884EA0",  "IDM"  = "#2494B5" ,  
            "IDF" = "#90C432" , "SDID" = "#D8609A")
names(Data0) = names(colors)
Up = list("IDM" = Data0$IDM$Up, "IDF" = Data0$IDF$Up)
Down = list("IDM" = Data0$IDM$Down, "IDF" = Data0$IDF$Down) 

X = if(length(Up)>1){RX_venn(rev(Up), colors, names= NULL,
                             title = "IDF vs IDM (logFC > 0)")} 
X
X = if(length(Down)>1){RX_venn(rev(Down), colors, name=NULL,
                               title = "IDF vs IDM (logFC < 0)")}
X
``` 

### SDIO {-} 

In this section we study the differences in the contrast 
SDIO based on the effect on female (green) and male (blue).

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
names(Datas) = Niv = contrasts2
Data = Datas$SDIO
Data <-  Data %>% 
    mutate(Significance = 
             case_when(
               
               logFC < 0 ~ "Down",
               logFC > 0 ~ "Up",
               TRUE ~ "-"
             ))
Data = Data[Data$p.adjust.BH < p_lim, ] %>% arrange(logFC)
IDs = Data$Symbol

ind = match(IDs, Datas$SDIO$Symbol)
Table = sapply(Datas, function(Dat){Dat[ind, c("logFC")]})
Table2 = sapply(Datas, function(Dat){Dat[ind, c("p.adjust.BH")] < 0.05})
Table = cbind(Table[,-1],Table2[,-1])
Table = setNames(data.frame(Table), c(Niv[-1], glue("{Niv[-1]}_FDR")))
DT = data.frame(Char  = factor(IDs, levels = IDs), Table)
```

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, fig.width=10}
#seq(from = 0.75, to = 19)
#seq(from = 1,25, to = 19)
max = max(unlist(Table))
min = min(unlist(Table))
SDIO = DT$SDIO

# Build the plot
pp <- (ggplot(DT) +
  geom_link(aes(y = IOF, yend = IOM,
                   x = Char, xend = Char, alpha = after_stat(index),
                size = after_stat(index)),  color = "#878787", size = 2) +
  geom_point(aes(y = IOM, x = Char), size = 5, color="#2494B5", alpha = 0.99) +
  geom_point(aes(y = IOF, x = Char), size = 5, color = "#90C432", alpha = 0.99) +   
  geom_point(shape = 21, aes(y = IOM, x = Char), size = 5, color="#2494B5") +
  geom_point(shape = 21,aes(y = IOF, x = Char), size = 5, color = "#90C432") +
  geom_hline(yintercept=0, linetype="dashed", 
                color = "#878787", size=0.01)  +
  geom_segment(aes(y = SDIO, yend = SDIO,
                   x = seq(from = 0.75, to = 19),
                   xend = seq(from = 1.25, to = 20)),
               color = "#CB326D", alpha = 0.5, size=1)  +
xlab(c("Genes")) +
ylab(c("LogFC")) +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))+

#hand-crafted legend
  theme(plot.background = element_rect(fill="transparent"),
            panel.background = element_rect(fill = "transparent"),
            axis.line = element_line(colour = "black"),
            axis.text.x = element_text(angle = 0, size = 8), legend.position = "none") ) 
pp
```



### Open Targets {-}

Comparison of the profiles obtained with those previously associated with 
obesity (blue), type 2 diabetes (green) or both, as reported in obesity (blue), 
type 2 diabetes (green) or both, according to information extracted from
extracted from Open Targets.  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
# Prepare data
Datas = sapply(files, function(x) get(load(file=x)), simplify = FALSE) 
 
Res = sapply(Datas, function(Data){
  Data <-  Data %>% 
    mutate(Significance = 
             case_when(
               p.adjust.BH > p_lim ~ "No",
               logFC < 0 ~ "Down",
               logFC > 0 ~ "Up",
               TRUE ~ "-"
             ))
  Data = Data %>% arrange((p.adjust.BH), abs(logFC)) 
  Up = Data[Data$Significance == "Up", "Symbol"]
  Down = Data[Data$Significance == "Down", "Symbol"]
  return(list(Up = Up, Down = Down))
}, simplify = FALSE)

# Lista relevantes
Relevantes = c("INSR", Res$SDIO$Up, Res$SDIO$Down, 
               "BCL9","ROR1","ARHGEF19","APCDD1L",
               "VAX2","TRABD2B","TMEM64","LRP4","NPHP3")
 
Res = sapply(Res,function(x)
  sapply(x, function(y)
    y[order(y %in% Relevantes, decreasing = T)], simplify = F), simplify = F) 

contrasts2 = Niv = c("IO", "IOM","IOF","SDIO")
names(Res) = contrasts2 

```   

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, results='hide' }
# Get table
Tabla = data.frame(Contrasts = names(Res),
  Up = rep("",4), Down = rep("", 4))
ft <- flextable(Tabla)
sapply(seq(names(Res)), function(i){
  sapply(seq(names(Res[[i]])), function(j){
    lis = Res[[i]][[j]]
    ft <<- eval(parse(text=glue('ft %>% mk_par(i={i}, j={j +1}, value = as_paragraph({paste({RXcolor(lis, 20)}, collapse=",")}))'))) 
  }, simplify = FALSE)
  }, simplify = FALSE)

BH = ft
```  


```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
Data=Tabla
header = data.frame(col_keys = colnames(Tabla),
                    H1 = c("Contrast", "Significance", "Significance"),
                    H2 = c("Contrast", "logFC > 0", "logFC < 0")
                    #H2=colnames(Data)
                    )

#Data = replace(Data, Data == 0, "-")
Color1 = "#007FA6"
Color2 = "#007FA6" 
Color.l = "black" 
Color.l2 = "#878787"
Color.up = "#CB326D"
Color.down = "#00AFBB" 
font = "Calibri"
footer = NULL
    
  
  ## Flextable
  # Set info
  # Set header
  ft <- set_header_df(ft, mapping = header, key = "col_keys" )
  # Format number
  ft <- colformat_num(ft, big.mark = ".", decimal.mark = "," )
  # Set footer
  if(! is.null(footer)){
    ft <- add_footer(ft, values = footer) %>% 
      #merge_h(part = "footer")  %>% 
      bold(j=1, part = "footer") %>%
      hline(border = fp_border(width = 2, color = Color1), part = "footer")
  }
  
  # Merge headers
  ft <- merge_h(ft, part = "header") %>% 
    merge_v( part = "header")
  
  # Merge Contrasts
  #ft <- merge_v(ft, j = c(1,2), part = "body")
  
  # Horizontal lines 
  ft <- hline(ft,border = fp_border(width = 2, color = Color1), part = "header")
  
  # Outer bottom
  ft <- hline_bottom(ft, border = fp_border(width = 2, color = Color1))
  
  # Alignment
  ft <- flextable::align(ft, align = "center", part = "header")
  ft <- flextable::align(ft, align = "center", part = "footer")
  ft <- flextable::align(ft, align = "center", part = "body")
  
  # First column aligned
  ft <- flextable::align(ft, j=1, align = "left", part = "all")
  
  
  # Header design
  x = Data[, 1]
    colors = case_when(
    x == "Up" ~ Color.up,
    x == "Down" ~ Color.down,
    TRUE ~ Color.l2
  )
  
  ft <- fontsize(ft, i = 1:2, size = 12, part = "header") %>% #fontsize(ft, i = 1:2, size = 12, part = "header") %>%
    color(i=1, color = Color2, part = "header")  %>%
    color(i=c(2), color = Color.l2, part = "header") %>%
    bold( bold = TRUE, part = "header") %>%
    bold( bold = TRUE, j=1, part = "body")  #%>%
    
    #color(j=1, color = colors, part = "body")
  # Rotate names
  #ft <- rotate(ft, i = 2, rotation = "btlr", part = "header", align = "bottom")
  
  # Font
  #
  
  # Color matrix
   ft<- fontsize(ft,  size = 12, part = "all") %>%
     font(fontname = font, part = "all")
ft <-ft %>% flextable::width(width = 0.3)

  # Footer
  if(! is.null(footer)){
    ft <- add_footer(ft, values = footer) %>% 
      merge_h(part = "footer")  %>% 
      bold(j=1, part = "footer") %>%
      hline(border = fp_border(width = 2, color = Color1), part = "footer")
      ft <- flextable::align(ft, align = "right", part = "footer")
  ft <- flextable::align(ft, j=1, align = "left", part = "footer")
  }

# We paint the values grey -
  colormatrix <- ifelse((as.matrix(Data)) %in% c(0,"0", "-"), Color.l2, Color.l )
ft <- color(ft, color = colormatrix, part = "body")  %>%
    color(j=1, color = "black", part = "body")
ft %>% width(width = 2)
```

```{r}
# saveRDS(ft, "./Data/OpenTargetsOb.rds")
```


## Diabetes {.tabset .tabset-fade -}   

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
DirData = "./Data/MA/Diabetes"

f= c("ID/Meta-analysis_ID_DF.RData", 
     "IDM/Meta-analysis_IDM_DF.RData",
     "IDF/Meta-analysis_IDF_DF.RData",
     "SDID/Meta-analysis_SDID_DF.RData")

files = glue("{DirData}/{f}")

```


### Resumen {-}

Resumen de los resultados obtenidos en el metaanálisis de diabetes de tipo 2.

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
# Prepare data
Datas = sapply(files, function(x) get(load(file=x)), simplify = FALSE)
p_lim = 0.05

Res = sapply(Datas, function(Data){
  Data <-  Data %>% 
    mutate(Significance = 
             case_when(
               logFC > 0 ~ "Up",
               logFC < 0 ~ "Down",
               TRUE ~ "-"
             ))
  res = table(Data[Data$p.adjust.BH < p_lim, "Significance"])
  return(c(res, Total = sum(res)))
}, simplify = FALSE)

contrasts2 = Niv = c("ID", "IDM","IDF","SDID")
names(Res) = contrasts2
```  


```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
Data = data.frame(Contrast = names(Res[[1]]),
                  Res, check.names = FALSE)
footer = c("Total number of genes", sapply(Datas, function(x) nrow(x)))
col_keys = colnames(Data)
names(footer) = col_keys
DataDM = Data
RX_table(Data, footer = footer) %>% width( width = 1.2, unit = "in")
```   


### Intersection {.tabset .tabset-fade -} 

Estudio de las intersecciones entre los genes obtenidos en el metaanalisis. 

#### Upset plot {-} 

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}
# Prepare data
p_lim = 0.05
logFC_lim = 0

# Gene lists
Direction = c("Up", "Down")

Data0 = sapply(Datas, function(Data){
  Up = rownames(Data[which(Data$p.adjust.BH < p_lim & Data$logFC > logFC_lim), ])
  Down = rownames(Data[which(Data$p.adjust.BH < p_lim & Data$logFC < 0), ])
  return(c("Up" = list(Up), "Down" = list(Down)))
}, simplify = FALSE)

names(Data0) = contrasts2 
```   


```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=10, fig.align = "center" }
MinSize = 0
An = unlist(Data0, recursive = FALSE) 
An = ComplexHeatmap::list_to_matrix(An)  

Nom = do.call(str_replace_all,
        c(list(colnames(An)), pattern = "\\.Up", replacement=" logFC>0"))
Nom = do.call(str_replace_all,
        c(list(Nom), pattern = "\\.Down", replacement=" logFC<0"))

Niv = c("ID, logFC>0", "ID, logFC<0", 
        "IDM, logFC>0", "IDM, logFC<0",
        "IDF, logFC>0", "IDF, logFC<0",
        "SDID, logFC>0", "SDID, logFC<0")
colnames(An) = Niv
Data = data.frame(An, check.names = FALSE)[,length(colnames(An)):1]

Direction = data.frame(set = colnames(Data),
                       #dir = rep(c("Up", "Down"), 4))
                       dir = rep(c("logFC > 0", "logFC < 0"), 4))

# Legend:
#Anots = as.data.frame(make_comb_mat(An)) 
#Leg = sapply(as.list(Anots), function(x) colnames(Data)[which(x == 1)], simplify = FALSE)
Legend = lapply(colnames(Data), function(x) upset_query(group= x,
                                                color="#878785",
                                                fill="#878787"))

#colnames(Data) = Names

UpsetPlot <- upset(Data, 
                   colnames(Data), 
                   name = "Intersecciones", guides = "over",
                   # Numero de interacciones minimas 
                   min_size =0,
    width_ratio=0.25,
    #stripes='transparent',
    # Sección horizontal
    stripes = upset_stripes(
      mapping=aes(color= dir),
      colors = c(
        "logFC > 0"= "grey89",#"#f2c4d6",
        "logFC < 0" = "white" #"#cbf0f2"
      ),
      data = Direction
    ), 
    matrix = intersection_matrix(
            geom=geom_point(
                shape=15,
                size=4,
                stroke = 1
            ), 
            segment=geom_segment(
              size= 1.2,
              arrow.fill = "#878787"
            ),
            outline_color=list(
                active='#878787',
                inactive='gray'
            )), #+ theme(text=element_text(size = 29,angle=0))
    
    # Colour of intersecting lines:
    # Assignment of colours
    #queries= Legend, 
    
    
    # We change the position of the set size
    set_sizes = upset_set_size(position='right',
                               #geom = geom_bar(color="red")
                               ) +    
      #geom_text(aes(label=..count..), #hjust=1.1,
      #         hjust=0, stat='count', color="white") + 
      ylab('Nº de genes significativos') + 
      theme(axis.text.x=element_text(size = 10,
                                     angle=0),
            text = element_text(size=10),
            panel.grid.minor = element_blank(),
            panel.grid.major = element_blank(),
            axis.line.x = element_line(color="#878787"),
            axis.ticks = element_line(color="#878787"),
            panel.background = element_rect(fill = "white",
                                          color = "white")),
    # Interactions
    base_annotations = list(
        'Intersection size'=( 
            intersection_size(
              mapping=aes(fill="bars_color"),
              alpha = 0.2,
              color = "#90C432", # color barra
              width = 0.6, # Ancho barra
              text = list(
                vjust=-0.5,
                #vjust=0.5,
                #hjust=-0.2,
                #angle=90,
                size =4.5
            ), 
            text_colors=c(
                on_background='#878787', on_bar="#878787"
            )
            ) + scale_fill_manual(values=c("bars_color"="#90C432"), guide="none") + # fill barra
              scale_alpha_manual(values = c(0.2))
            + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(),
               plot.background = element_rect(fill="transparent",
                                                   color = "transparent"),
          panel.background = element_rect(fill = "transparent",
                                          color = "transparent"),
          axis.line.y = element_line(color="#878787"),
          axis.text.y=element_text(size = 10, angle=0)
          #axis.line.x = element_line(color="black"),
          #axis.ticks = element_line(color="black")
          )
            + ylab('Genes significativos por grupo')
        )),
    sort_sets = FALSE
        
)  

UpsetPlot 

```

#### Venn plot {-} 

```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, fig.width=4, fig.align='center'}
# Up
colors = c( "ID" = "#884EA0",  "IDM"  = "#2494B5" ,  
            "IDF" = "#90C432" , "SDID" = "#D8609A")
names(Data0) = names(colors)
Up = list("IDM" = Data0$IDM$Up, "IDF" = Data0$IDF$Up)
Down = list("IDM" = Data0$IDM$Down, "IDF" = Data0$IDF$Down) 

X = if(length(Up)>1){RX_venn(rev(Up), colors, names= NULL,
                             title = "IDF vs IDM (logFC > 0)")} 
X
X = if(length(Down)>1){RX_venn(rev(Down), colors, name=NULL,
                               title = "IDF vs IDM (logFC > 0)")} 
X
``` 



### Open Targets {-}  

Comparison of the profiles obtained with those previously associated with 
obesity (blue), type 2 diabetes (green) or both, according to information
extracted from Open Targets.

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
# Prepare data
Datas = sapply(files, function(x) get(load(file=x)), simplify = FALSE) 
 
Res = sapply(Datas, function(Data){
  Data <-  Data %>% 
    mutate(Significance = 
             case_when(
               p.adjust.BH > p_lim ~ "No",
               logFC < 0 ~ "Down",
               logFC > 0 ~ "Up",
               TRUE ~ "-"
             ))
  Data = Data %>% arrange((p.adjust.BH), abs(logFC))  
  Up = Data[Data$Significance == "Up", "Symbol"]
  Down = Data[Data$Significance == "Down", "Symbol"]
  return(list(Up = RX_ifelse(length(Up) == 0, "_", Up ), Down = RX_ifelse(length(Down) == 0, "_", Down )))
}, simplify = FALSE)
contrasts2 = Niv = c("ID", "IDM","IDF","SDID") 
names(Res) = contrasts2

Res[c("IDM", "IDF")] = sapply(Res[c("IDM", "IDF")],function(x)
  sapply(x, function(y)
    y[order(y %in% Relevantes, decreasing = T)], simplify = F), simplify = F) 
```   

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, results='hide' }
# Get table
Tabla = data.frame(Contrasts = names(Res),
  Up = rep("",4), Down = rep("", 4))
ft <- flextable(Tabla)
sapply(seq(names(Res)), function(i){
  sapply(seq(names(Res[[i]])), function(j){
    lis = Res[[i]][[j]]
    ft <<- eval(parse(text=glue('ft %>% mk_par(i={i}, j={j +1}, value = as_paragraph({paste({RXcolor(lis, 20)}, collapse=",")}))'))) 
  }, simplify = FALSE)
  }, simplify = FALSE)

BH = ft
```  


```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
Data=Tabla
header = data.frame(col_keys = colnames(Tabla),
                    H1 = c("Contrast", "Significance", "Significance"),
                    H2 = c("Contrast", "logFC > 0", "logFC < 0")
                    #H2=colnames(Data)
                    )

#Data = replace(Data, Data == 0, "-")
Color1 = "#007FA6"
Color2 = "#007FA6" 
Color.l = "black" 
Color.l2 = "#878787"
Color.up = "#CB326D"
Color.down = "#00AFBB" 
font = "Calibri"
footer = NULL
    
  
  ## Flextable
  # Set info
  # Set header
  ft <- set_header_df(ft, mapping = header, key = "col_keys" )
  # Format number
  ft <- colformat_num(ft, big.mark = ".", decimal.mark = "," )
  # Set footer
  if(! is.null(footer)){
    ft <- add_footer(ft, values = footer) %>% 
      #merge_h(part = "footer")  %>% 
      bold(j=1, part = "footer") %>%
      hline(border = fp_border(width = 2, color = Color1), part = "footer")
  }
  
  # Merge headers
  ft <- merge_h(ft, part = "header") %>% 
    merge_v( part = "header")
  
  # Merge Contrasts
  #ft <- merge_v(ft, j = c(1,2), part = "body")
  
  # Horizontal lines 
  ft <- hline(ft,border = fp_border(width = 2, color = Color1), part = "header")
  
  # Outer bottom
  ft <- hline_bottom(ft, border = fp_border(width = 2, color = Color1))
  
  # Alignment
  ft <- flextable::align(ft, align = "center", part = "header")
  ft <- flextable::align(ft, align = "center", part = "footer")
  ft <- flextable::align(ft, align = "center", part = "body")
  
  # First column aligned
  ft <- flextable::align(ft, j=1, align = "left", part = "all")
  
  
  # Header design
  x = Data[, 1]
    colors = case_when(
    x == "Up" ~ Color.up,
    x == "Down" ~ Color.down,
    TRUE ~ Color.l2
  )
  
  ft <- fontsize(ft, i = 1:2, size = 12, part = "header") %>% #fontsize(ft, i = 1:2, size = 12, part = "header") %>%
    color(i=1, color = Color2, part = "header")  %>%
    color(i=c(2), color = Color.l2, part = "header") %>%
    bold( bold = TRUE, part = "header") %>%
    bold( bold = TRUE, j=1, part = "body")  #%>%
    
    #color(j=1, color = colors, part = "body")
  # Rotate names
  #ft <- rotate(ft, i = 2, rotation = "btlr", part = "header", align = "bottom")
  
  # Font
  #
  
  # Color matrix
   ft<- fontsize(ft,  size = 12, part = "all") %>%
     font(fontname = font, part = "all")
ft <-ft %>% flextable::width(width = 0.3)

  # Footer
  if(! is.null(footer)){
    ft <- add_footer(ft, values = footer) %>% 
      merge_h(part = "footer")  %>% 
      bold(j=1, part = "footer") %>%
      hline(border = fp_border(width = 2, color = Color1), part = "footer")
      ft <- flextable::align(ft, align = "right", part = "footer")
  ft <- flextable::align(ft, j=1, align = "left", part = "footer")
  }

# We paint the values grey -
  colormatrix <- ifelse((as.matrix(Data)) %in% c(0,"0", "-"), Color.l2, Color.l )
ft <- color(ft, color = colormatrix, part = "body")  %>%
    color(j=1, color = "black", part = "body")
ft %>% width(width = 2)
```   

```{r}
# saveRDS(ft, "./Data/OpenTargetsT2D.rds")
```


***      

&nbsp;  

