---
title: "Meta-analysis"
author: "Roxana Andreea Moldovan Moldovan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_float: false
    number_sections: false 
    code_folding: hide
    theme: flatly
    bg: '#FFFFFF'
    fg: '#202020'
    primary: '#90C432'
    secondary: '#2494b5' 
  pdf_document:
    extra_dependencies: "subfig"
linkcolor: blue
urlcolor: blue
citecolor: blue
header-includes:
- \usepackage{subfig}
- \usepackage{float}
fig_caption: yes 
link-citations: TRUE
---

```{r}
# ~~~~~~~~~~~~ ~~~~~~~~~~~~ ~~~~~~~~~ ** ~~~~~~~~~ ~~~~~~~~~~~~ ~~~~~~~~~~~~ #


# ~~~~~~~~~~~~ Loading packages ~~~~~~~~~~~~ #

library(pacman)
pacman::p_load(argparse)
pacman::p_load(glue)
pacman::p_load(limma)
pacman::p_load(metafor)
pacman::p_load(parallel)
pacman::p_load(stringr)


# ~~~~~~~~~~~~ Functions ~~~~~~~~~~~~ #

source("../Functions/Functions.R")

RX_getSE <- function(Data # Fit limma
){
  # Método 1:
  TopTab <- topTable(Data, number = "all", confint=TRUE, adjust.method = "fdr", sort.by = "none")
  # Calculamos el SE
  TopTab[, "SE"] <- (TopTab[, "CI.R"] - TopTab[, "CI.L"])/ 3.92
  
  # Método 2:
  coefSE = sqrt(Data$s2.post) * Data$stdev.unscaled
  TopTab[, c("coef", "coefSE")] = c(Data$coefficients, coefSE)
  
  return(TopTab)
}

RX_MetaAnalysis_prep <- function(Studies, # Studies accesion to include
                                 Tissue = c("SAT", "VAT", "LSAT"), # One or several tissues to include in the meta-analysis to include
                                 Contrast, # Contrast to use
                                 Datas, # data
                                 SE_coef = c("SE", "coefSE") # SE to take
                                 ){
  # Check if the arguments are suitable
  match.arg(Tissue, several.ok = TRUE)
  #match.arg(Disease, several.ok = FALSE)
  match.arg(SE_coef, several.ok = FALSE)
  
  # Extract the information from the indicated tissues and contrast
  Data0 = sapply(Studies,
                 function(st) sapply(Tissue,
                                     function(ts) sapply(Contrast,
                                                         function(c) Datas[[st]][[ts]][[c]], 
                                                         simplify = FALSE),
                                     simplify = FALSE),
                 simplify = FALSE)
  
  Data = unlist(unlist(Data0, recursive = FALSE), recursive = FALSE)
  names(Data) = Studies
  Data = Filter(Negate(is.null), Data)
  
  #------------- 1. Preparing input for meta-analysis
  # Get SE
  DataSE = sapply(names(Data), function(x) RX_getSE(Data[[x]]), simplify = FALSE) 
  # Get Gene and annotation
  IDs = unlist(sapply(DataSE, function(x) rownames(x), simplify = TRUE), use.names = FALSE)
  SYMBOL = sapply(Data, function(x) x$TopTab[,c("ENTREZID", "ENSEMBL", "SYMBOL")], simplify = FALSE)
  ANOT = do.call("rbind", SYMBOL)
  cat(glue("\n{length(unique(IDs))} of the {length(IDs)} genes have been taken."), "\n") 
  IDs = unique(IDs)
  ANOT = ANOT[match(unique(IDs), ANOT$ENTREZID),]
  rownames(ANOT) = ANOT$ENTREZID
  
  # LogFC matrix
  MatLogFC <- sapply(DataSE, function(x) x[IDs, "logFC"]) 
  rownames(MatLogFC) <- IDs
  
  # SD matrix
  MatSE <- sapply(DataSE, function(x) x[IDs, SE_coef]) 
  rownames(MatSE) <- IDs
  
  # Filter keeping the genes that appear in more than one study
  keep = which(rowSums(!is.na(MatSE)) > 1)
  MatLogFC = MatLogFC[keep,]
  MatSE = MatSE[keep,] 
  
  return(list("MatLogFC" = MatLogFC, "MatSE" = MatSE, Anot = ANOT))
}

```

```{r}
#------------- Checking arguments 
getwd()
### XX NONO
args = list()
args$studies = "ALL"
args$tissue = "SAT"
args$filein = "C:/Users/roxya/OneDrive/Documentos/01Master_bioinformatica/00TFM/Git/T2D-Meta-Analysis/Data/DifferentialExpressionObesity.RData"
args$outdir = "C:/Users/roxya/OneDrive/Documentos/01Master_bioinformatica/00TFM/Git/T2D-Meta-Analysis/Data/Obesity"
args$prefix = "Meta-analysis"
#args$contrast = "Ob - C, Ob.M - C.M, Ob.F - C.F, (Ob.M - C.M) - (Ob.F - C.F)"
args$contrast = "Ob - C"
args$report = TRUE
args$redo = FALSE
args$plim = 0.05


#load(file="../Data/DE/DifferentialExpressionObesity.RData")
#Rscript 05MetaAnalysis.R -t SAT -f "../Data/DiffExprsObesity.RData" -o ../Data/Meta-analysis_Obesity_all.RData -c "Ob - C, Ob.M - C.M, Ob.F - C.F, (Ob.M - C.M) - (Ob.F - C.F)"


# Load the data
Datas = get(load(args$filein))

# Get the studies
if (args$studies != "ALL"){
  # Take studies to include
  Studies = stringr::str_split_1(args$studies, pattern = ",")
  Studies = trimws(Studies, whitespace = "[ \t\r\n\\.]")
} else{
  Studies = names(Datas)
}

# Check
if (! all(Studies %in% names(Datas))){
  cat("The indicated file does not contain information on",
      "the following studies. \n")
  cat("\t", Studies[! Studies %in% names(Datas)])
  cat("\n")
  stop("Please review input.",call. = FALSE)
}

# Get the tissues
Tissue = stringr::str_split_1(args$tissue, pattern = ",")
Tissue = trimws(Tissue, whitespace = "[ \t\r\n\\.]")

# Get the contrasts
Contrasts = stringr::str_split_1(args$contrast, pattern = ",")
Contrasts = trimws(Contrasts, whitespace = "[ \t\r\n\\.]")

method = "DL"

PlotsDir = glue("{args$outdir}/Plots")

# Output directory
if(! file.exists(args$outdir)){
  system(glue("mkdir {args$outdir}"))
}
# Plots directory
system(glue("mkdir PlotsDir"))
```

```{r}
#------------- 1. Meta-analysis for genes
cat("\n---------------------------------------------\n")
cat("Analyzing...\n") 
contrast = Contrasts[[1]]
Mats_out = glue("{args$outdir}/{args$prefix}_mats.RData")
MA_out = glue("{args$outdir}/{args$prefix}_MA.RData")
DF_out = glue("{args$outdir}/{args$prefix}_DF.RData")

if(file.exists(Mats_out) & isFALSE(args$redo)){
  # Load previous data
  load(file=Mats_out)
}else{
  Mats = RX_MetaAnalysis_prep(Studies = Studies,
                            Tissue = Tissue,
                            Datas = Datas,
                            SE_coef = "coefSE",
                            Contrast = contrast)
  save(Mats, file = Mats_out)
}

if(file.exists(MA_out) & isFALSE(args$redo)){
  # Load previous data
  load(file=MA_out)
}else{
  #------------- 1. Meta-analysis for genes  
  MA <- sapply(rownames(Mats$MatLogFC),
               function(x) metafor::rma(yi = Mats$MatLogFC[x, ],
                                        sei = Mats$MatSE[x, ],
                                        method = method,
                                        verbose = FALSE), simplify = FALSE, USE.NAMES = TRUE ) 
  save(MA, file = MA_out)
}

if(file.exists(DF_out) & isFALSE(args$redo)){
  # Load previous data
  load(file=DF_out)
}else{
  # Data frame including the detailed results.
  resultMA <- as.data.frame(do.call("rbind",
                                    sapply(MA,
                                           function(x){c("interv_inf" = x$ci.lb, 
                                                         "Summary_logFC" = x$b, 
                                                         "interv_sup" = x$ci.ub,
                                                         "pvalue" = x$pval,
                                                         x$QE,x$QEp, x$se,
                                                         x$tau2, x$I2, x$H2)},
                                           simplify = FALSE)))
  # Adjust p.values
  resultMA$p.adjust.fdr <- stats::p.adjust(resultMA$pvalue, method = "fdr")
  resultMA$p.adjust.BY  <- stats::p.adjust(resultMA$pvalue, method = "BY")
  resultMA$logFDR <- -log10(resultMA$p.adjust.fdr)
  resultMA$N <- rowSums(!is.na(Mats$MatSE))
  save(resultMA, file = DF_out)
}
Datas$E_MEXP_1425$SAT$`Ob - C`$TopTab


```

```{r}
args$plim = 0.00000000001
sigData = resultMA[resultMA[, "p.adjust.fdr"] < args$plim,]
sigData = sigData[order(abs(sigData[, "p.adjust.fdr"])),]
color = "#2494b5"
for (gene in rownames(sigData)){
  res <- MA[[gene]]
  res <- rma(yi= Mats$MatLogFC[mygenes,], sei =Mats$MatSE[mygenes,], method = "DL",
      slab = Studies)
  # Forest plot
  svg(glue("{PlotsDir}/{args$prefix}_{gene}forest.svg"))
  forest(res, 
         slab = toupper(colnames(Mats$MatLogFC)),
         xlab="Fold Change", cex=0.7,
         mlab="Modelo 'DL' para todos los estudios.", col = color, 
         main = paste("\n", Mats$Anot[gene, "SYMBOL"] , " (", gene, ")",sep=""))    
  text( 9,-3, "logFC [IC 95%]", pos=2, cex = 0.7)
  dev.off()
  
  #Funnel plot
  svg(glue("{PlotsDir}/{args$prefix}_{gene}funnel.svg"))
  funnel(res, main=paste("\n", Mats$Anot[gene, "SYMBOL"] , " (", gene, ")",sep=""), back ="transparent", ylab = "Error estándar",
         xlab = "logFC", label = "out", shade=c("#F0EFEF", "#90C432", "#2494b5"), refline=0,
         level=c(90, 95, 99), legend=TRUE, atransf=exp)
  dev.off()  
  
  #Influence plot
  svg(glue("{PlotsDir}/{args$prefix}_{gene}influence.svg"))
  inf <- influence(res)
  plot(inf)
  dev.off()  
  
}


```

```{r}
report_out = glue("{args$outdir}/{args$prefix}Report.rmd")
if(isTRUE(args$report)){
    # File info
cat('---
title: "05 Metaanálisis."
author: "Roxana Andreea Moldovan Moldovan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_float: false
    number_sections: false 
    code_folding: hide
    theme: flatly
    bg: "#FFFFFF"
    fg: "#202020"
    primary: "#90C432"
    secondary: "#2494b5" 
  pDT_document:
    extra_dependencies: "subfig"
linkcolor: blue
urlcolor: blue
citecolor: blue
header-includes:
- \\usepackage{subfig}
- \\usepackage{float}
fig_caption: yes 
link-citations: TRUE 
---
        ',
  sep ="",
  file = report_out,
  append = FALSE
  )

  ## Meta analisis
cat('\n## Meta-análisis {.tabset .tabset-fade -} \n',
    '&nbsp;  \n\n',
    sep ="",
    file = report_out,
    append = TRUE)

# Load packages
cat(
  # Open chunk
  '\n```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}\n',
  '# Load packages \n',
  'library(pacman)
pacman::p_load(glue)
pacman::p_load(DT)
pacman::p_load(magick)',
  # Close chunk
'\n```  \n\n',
sep ="",
  file = report_out,
  append = TRUE)

cat('\nEn este documento se muestran los resultados del metaanálisis de los estudios: ', 
    paste(Studies, collapse = ", "),
    ' para el contraste ',
    args$contrast, 
    '.  \n&nbsp;  \n\n',
    sep ="",
    file = report_out,
    append = TRUE)

cat('\n### Tabla de resultados {-} \n',
    '&nbsp;  \n\n',
    sep ="",
    file = report_out,
    append = TRUE)

cat(
  # Open chunk
  '\n```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}\n',
  'DT = get(load("', DF_out, '"))\n',
  'datatable(DT, options = list(caption = "Resultados del metaanálisis.",
                              scrollX = TRUE))',
  # Close chunk
'\n```  \n\n','\n&nbsp;  \n\n',
sep ="",
  file = report_out,
  append = TRUE)

cat('\n### Representación gráfica {-}  \n&nbsp; \n',
    '\nAquí se muestran las figuras características del metaanálisis para los
genes significativos tomando un p valor ajustado por BH < ', args$plim, '.  \n',
    '&nbsp;  \n\n',
    sep ="",
    file = report_out,
    append = TRUE)

# Function
cat(
  # Open chunk
  '\n```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE}\n',
  '# Función para repesentar las figuras\n',
  'grid_img <- function(list_fig, dir, patt, dim = 300){
  # Abrimos todas las filas:
  forest_plots = c()
  funnel_plots = c()
  influ_plots = c()
  for (fig in list_fig){
    f1 = glue("{fig}forest.svg")
    out1 <- magick::image_read(paste0(dir, "/", f1))
    forest_plots = c(forest_plots, image_scale(out1, dim))
    
    f2 = glue("{fig}funnel.svg")
    out2 <- magick::image_read(paste0(dir,"/", f2))
    funnel_plots = c(funnel_plots, image_scale(out2, dim))  
    
    f3 = glue("{fig}influence.svg")
    out3 <- magick::image_read(paste0(dir,"/", f3))
    influ_plots = c(influ_plots, image_scale(out3, dim))
  } 
  
  lim = length(forest_plots)
  grid = NULL
  for ( i in seq(list_fig)){
    a = forest_plots[[i]]
    b = funnel_plots[[i]]
    c = influ_plots[[i]]
    
    fila <- image_append(c(a, b, c))
    
    if(is.null(grid)){
      grid = fila
    }else{
      grid = image_append(c(grid,fila),  stack = TRUE)
    } 
  }
    return(grid)
}',
  # Close chunk
'\n```  \n\n','\n&nbsp;  \n\n',
sep ="",
  file = report_out,
  append = TRUE)
}

cat(
  # Open chunk
  '\n```{r echo=TRUE, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE, out.width=".99\\\\linewidth"}\n',
  'path = "',PlotsDir, '"\n',
  'files <- list.files(path = path, pattern = "',glue("{args$prefix}.*forest"),'")\n',
  'figuras = sapply( files, function(x) substr(x, start = 0, stop = nchar(x)-10), 
                  USE.NAMES = FALSE)\n',
  '# Plot\n',
  'grid_img(figuras, dir = path)',
  # Close chunk
'\n```  \n\n','\n&nbsp;  \n\n',
sep ="",
  file = report_out,
  append = TRUE)


# Create HTML
rmarkdown::render(report_out)
```

```{r}

```



```{r}
library(magick)
grid_img <- function(list_fig, dir, patt){
  # Abrimos todas las filas:
  forest_plots = c()
  funnel_plots = c()
  influ_plots = c()
  for (fig in list_fig){
    f1 = glue("{fig}forest.svg")
    out1 <- magick::image_read(paste0(dir, "/", f1))
    forest_plots = c(forest_plots, image_scale(out1, 225))
    
    f2 = glue("{fig}funnel.svg")
    out2 <- magick::image_read(paste0(dir,"/", f2))
    funnel_plots = c(funnel_plots, image_scale(out2, 225))  
    
    f3 = glue("{fig}influence.svg")
    out3 <- magick::image_read(paste0(dir,"/", f3))
    influ_plots = c(influ_plots, image_scale(out3, 225))
  } 
  
  lim = length(forest_plots)
  grid = NULL
  for ( i in seq(list_fig)){
    a = forest_plots[[i]]
    b = funnel_plots[[i]]
    c = influ_plots[[i]]
    
    fila <- image_append(c(a, b, c))
    
    if(is.null(grid)){
      grid = fila
    }else{
      grid = image_append(c(grid,fila),  stack = TRUE)
    } 
  }
    return(grid)
}
```

```{r}
glue("{PlotsDir}/{args$prefix}_{gene}")
PlotsDir
files <- list.files(path = PlotsDir, pattern = glue("{args$prefix}.*forest")) 
figuras = sapply( files, function(x) substr(x, start = 0, stop = nchar(x)-10), 
                  USE.NAMES = FALSE)  
grid_img(figuras, dir = PlotsDir)
```


```{r}
sig.results <- resultMA[resultMA[, "p.adjust.fdr"] < 0.05,]
sig.results = sig.results[order(sig.results[, "p.adjust.BY"]),]

for (i in 1){#seq(nrow(sig.results))){
  mygenes <- rownames(sig.results)[i]
  res <- MA[[mygenes]]
  #FOREST PLOT
  # png (filename = paste("FOREST_", mygenes,".png", sep =""), width = 960 ,
  #      height = 960, res = 200)
  'png (filename = paste(gsub("-","_",mygenes),"_FOREST",".png", sep =""), width = 960 ,
       height = 960, res = 200)'
  a = forest(res,
         slab = toupper(colnames(Mats$MatLogFC)), #Nombre de los estudios
         xlab="logFC", cex=0.7,
         mlab=paste(method, "Model for All Studies", sep = " "),
         border = "black", #Color del borde del rombo
         col = "red", #Color del rombo
         main = paste("\n", mygenes, sep=""))
  text( 9,-3, "logFC [IC 95%]", pos=2, cex = 0.7)
  'dev.off()'
  
  #FUNNEL PLOT
  'png (filename = paste(gsub("-","_",mygenes),"_FUNNEL", ".png", sep =""), width = 960 ,
       height = 960, res = 200)'
  par(mfrow=c(2,2))
  funnel(res, main="Standard Error", back ="darkslategray1",
         xlab = paste("logFC (", mygenes, ")",sep =""))
  funnel(res, yaxis="vi", main="Sampling Variance", back ="darkslategray1",
         xlab = paste("logFC (", mygenes, ")",sep =""))
  funnel(res, yaxis="seinv", main="Inverse Standard Error",
         back ="darkslategray1", xlab = paste("logFC (", mygenes, ")",sep =""))
  funnel(res, yaxis="vinv", main="Inverse Sampling Variance",
         back ="darkslategray1",  xlab = paste("logFC (", mygenes, ")",sep =""))
  par(mfrow=c(1,1))
  'dev.off() '
  
   #INFLUENCE PLOTS
  # That shows various diagnostic measures
  'png (filename = paste(gsub("-","_",mygenes), "_INFLUENCE", ".png", sep =""), width = 960 ,
       height = 960, res = 200) ##CAMBIAR'
  inf <- influence(res)
  #plot(inf, plotfb = T)#"plotfb" is not a graphical parameter
  plot(inf)
  'dev.off()'
}

res
```

```{r}
  forest(res, 
         slab = toupper(colnames(Mats$MatLogFC)),
         xlab="Fold Change", cex=0.7,
         mlab=paste("DL", "Model for All Studies", sep = " "), col = "red", 
         main = paste("\n", Mats$Anot[gene, "SYMBOL"] , " (", gene, ")",sep=""))    
  text( 9,-3, "Odds Ratio [IC 95%]", pos=2, cex = 0.7)
```


```{r}
library(metaviz)
viz_forest(x = res, variant = "classic")
```
```{r}
library(ggplot2)
a = viz_funnel(res, sig_contours = FALSE,
               addev_contours = FALSE,
               trim_and_fill = FALSE,
               egger = TRUE,
               contours_col = "Greens") 
a + scale_fill_manual(values=c("red", "red", "green", "black", "purple", "red", "purple"))
a$scales
layer_scales(a)$y$range$range
x = abs(layer_scales(a)$y$range$range[[1]])

a  + coord_cartesian(ylim = c(x-0.05*x,-0.01)) 

homeopath[, c("d", "se")]
age <- factor(Studies, labels = Studies)
```

```{r}
class(funnel(res))
```


```{r}
# NO
sig.genes.df = resultMA[which(resultMA[, "p.adjust.fdr"] < 0.05),]
#sig.genes.df = sig.results
for (i in 1){#rownames(sig.genes.df)){
  #define studies for each function (not NA)
  estudios <- stringr::str_split_i(colnames(Mats$MatSE)[!is.na(Mats$MatSE)[i,]], pattern = "\\.", i=1)

  #influence info 1:
  #number of studies where the sign of the logOR is the same  of the global logOR:
  sig.genes.df[i, "infl.same.sign.logFC"] <- sum(sign(MA[[i]]$yi)== rep(sign(MA[[i]]$b),length(estudios)))

  #influence info 2: how many studies could be influencers?
  inf <- influence(MA[[i]])
  res <- paste(estudios[inf$is.infl], collapse = ",")
  sig.genes.df[i, "infl.nstudies"] <- ifelse(res =="", "non", res)

  #sensivity analysis
  l1 <-as.data.frame(leave1out(MA[[i]]))
  rownames(l1) <- estudios

  #1. p.value about differences between all estimates from leave one out
  #   and global estimate)
  sig.genes.df[i, "sensi.global"] <-t.test(x= l1$estimate,
                                           mu=as.numeric(MA[[i]]$b))$p.value
  #2. number of  studies where pvalue > 0.05
  # (we hope p-values < 0.05, significant estimates)
  res2 <- paste(estudios[l1$pval > 0.05], collapse = ",")
  sig.genes.df[i, "sensi.specific"] <- ifelse(res2 =="", "all.p.values < 0.05", res2)
}
```

```{r}
sig.results <- resultMA[resultMA[, "p.adjust.fdr"] < 0.05,]
sig.results = sig.results[order(sig.results[, "p.adjust.BY"]),]

for (i in 1){#seq(nrow(sig.results))){
  mygenes <- rownames(sig.results)[i]
  res <- MA[[mygenes]]
  #FOREST PLOT
  # png (filename = paste("FOREST_", mygenes,".png", sep =""), width = 960 ,
  #      height = 960, res = 200)
  'png (filename = paste(gsub("-","_",mygenes),"_FOREST",".png", sep =""), width = 960 ,
       height = 960, res = 200)'
  forest(res,
         slab = toupper(colnames(Mats$MatLogFC)), #Nombre de los estudios
         xlab="logFC", cex=0.7,
         mlab=paste(method, "Model for All Studies", sep = " "),
         border = "black", #Color del borde del rombo
         col = "red", #Color del rombo
         main = paste("\n", mygenes, sep=""))
  text( 9,-3, "logFC [IC 95%]", pos=2, cex = 0.7)
  'dev.off()'
  
  #FUNNEL PLOT
  'png (filename = paste(gsub("-","_",mygenes),"_FUNNEL", ".png", sep =""), width = 960 ,
       height = 960, res = 200)'
  par(mfrow=c(2,2))
  funnel(res, main="Standard Error", back ="darkslategray1",
         xlab = paste("logFC (", mygenes, ")",sep =""))
  funnel(res, yaxis="vi", main="Sampling Variance", back ="darkslategray1",
         xlab = paste("logFC (", mygenes, ")",sep =""))
  funnel(res, yaxis="seinv", main="Inverse Standard Error",
         back ="darkslategray1", xlab = paste("logFC (", mygenes, ")",sep =""))
  funnel(res, yaxis="vinv", main="Inverse Sampling Variance",
         back ="darkslategray1",  xlab = paste("logFC (", mygenes, ")",sep =""))
  par(mfrow=c(1,1))
  'dev.off() '
  
   #INFLUENCE PLOTS
  # That shows various diagnostic measures
  'png (filename = paste(gsub("-","_",mygenes), "_INFLUENCE", ".png", sep =""), width = 960 ,
       height = 960, res = 200) ##CAMBIAR'
  inf <- influence(res)
  #plot(inf, plotfb = T)#"plotfb" is not a graphical parameter
  plot(inf)
  'dev.off()'
}

res
```



```{r}

```

