---
title: "06FunctionalAnalysisReport"
author: "Roxana Andreea Moldovan Moldovan"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: false
    toc_float: false
    number_sections: false 
    code_folding: hide
    theme: flatly
    bg: '#FFFFFF'
    fg: '#202020'
    primary: '#90C432'
    secondary: '#2494b5' 
  pdf_document:
    extra_dependencies: "subfig"
linkcolor: blue
urlcolor: blue
citecolor: blue
header-includes:
- \usepackage{subfig}
- \usepackage{float}
fig_caption: yes 
link-citations: TRUE

---  

```{r setup, include=FALSE}    
loadData = FALSE
preloadData = TRUE 
```

# GSEA summary {.tabset .tabset-fade -}

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, results = "hide"}
# Packages
pacman::p_load(ComplexHeatmap)
pacman::p_load(ComplexUpset)
pacman::p_load(dplyr)
pacman::p_load(flextable)
pacman::p_load(ggforce)
pacman::p_load(ggplot2)
pacman::p_load(ggVennDiagram)
pacman::p_load(glue)
pacman::p_load(Hmisc)
pacman::p_load(officer)
pacman::p_load(rrvgo)
pacman::p_load(sf)
pacman::p_load(stringr)  

#Params
DirBase = "./"
DirData = glue("{DirBase}/Data_no")
DirOut = glue("{DirData}/FA") 
```  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
# Functions
source("../Functions/Functions.R")

RX_revigo <- function(reducedTerms,
                      wrap = 11){
  
  plot_df = data.frame(sort(table((reducedTerms$parentTerm)),
                            decreasing = T))[c(0:20),]
  colnames(plot_df) = c("region", "n")
  plot_df$region = factor(str_to_sentence(plot_df$region), levels=str_to_sentence(plot_df$region))
  plot_df$ord = match(plot_df$n,unique(plot_df$n)) 
  
  maxy = log2(max(plot_df$n) + max(plot_df$n)*0.5)
  plt <- ggplot(plot_df) +
    # Make custom panel grid
    geom_hline(
      aes(yintercept = y), 
      data.frame(y = c(0,cumsum(rep(maxy/3, 3)))),
      color = "lightgrey"
    ) + 
    # Add bars to represent the cumulative track lengths
    # str_wrap(region, 5) wraps the text so each line has at most 5 characters
    # (but it doesn't break long words!)
    geom_col(
      aes(
        x = reorder(str_wrap(region, wrap), n), 
        y = log2(n),
        fill = ord
      ),
      position = "dodge2",
      show.legend = TRUE,
      alpha = .9
    ) + 
    
    # Lollipop shaft for mean gain per region
    geom_segment(
      aes(
        x = reorder(str_wrap(region, wrap), n),
        y = 0,
        xend = reorder(str_wrap(region, wrap), n),
        yend = (maxy + maxy*0.1)
      ),
      linetype = "dashed",
      color = "gray25",
    )  +
    geom_text(
      aes(
        x = reorder(str_wrap(region, wrap), n),
        y = log2(n) + 0.1,
        label = n,
        angle = 360 + (360 * (rank(-(n)) - 0.5) / max(rank(-n))) #360 - (360 * (rank(-seq(n))-0.5)) /length(plot_df$n), 
      ),
      position = position_dodge(width = 1),
      vjust = 0, size = 4, color ="#878787")+ 
    
    # Make it circular!
    coord_polar()+ 
  
  
    # New fill and legend title for number of tracks per region
    scale_fill_gradientn(
      "Amount of Tracks",
       colours = rev(c("#cef587", "#90C432", "#2494b5","#03556e"  ))
    ) +
    # Make the guide for the fill discrete
    guides(
      fill = "none"
    ) +
    theme(
      # Remove axis ticks and text
      axis.title = element_blank(),
      axis.ticks = element_blank(),
      axis.text.y = element_blank(),
      # Use gray text for the region names
      axis.text.x = element_text(color = "#878787", size = wrap),
      # Move the legend to the bottom
      legend.position = "bottom",
      #panel.background = element_rect(fill = "white", color = "white"),
      panel.grid = element_blank(),
      panel.grid.major.x = element_blank(),
      panel.background = element_rect(fill = "transparent", color = NA)
    )
  
  
  return(plt)
}
```


## Summary {-}   

The GSEA results are summarised below:  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE}
ID = c("BP", "CC", "MF", "KEGG", "Reactome")

# Obesity
Niv = c("IO", "IOM", "IOF", "SDIO" )
Datasob = sapply(ID, function(x){
  setNames(sapply(get(load(glue("{DirOut}/Obesity_gse{x}.RData"))), function(y)
    data.frame(y), simplify = FALSE),Niv)
}, simplify=FALSE)
Countob = sapply(Datasob, function(x) sapply(x, nrow))
DFOB = t(Countob)

# Diabetes
Niv = c("ID", "IDM", "IDF", "SDID" )
DatasDM = sapply(ID, function(x){
  setNames(sapply(get(load(glue("{DirOut}/FA_gse{x}.RData"))), function(y)
    data.frame(y), simplify = FALSE),Niv)
}, simplify=FALSE)
CountDM = sapply(DatasDM, function(x) sapply(x, nrow))
DFDM = t(CountDM)

# Table:
Df = cbind(DFOB, DFDM)

```


```{r echo=T, eval= TRUE, message=FALSE, warning=FALSE, fig.width=12, fig.align = "center" }
## Representation
# Data:
Data = data.frame(DB = c("GO: Biological processes",
                         "GO: Cellular components",
                         "GO: Molecular functions",
                         "KEGG Pathways",
                         "Reactome Pathways"),
                  (Df))
header = data.frame(col_keys = colnames(Data),
                    H1 = c("DB", rep("Obesity", 4),rep("Diabetes", 4)),
                    H2=colnames(Data))

Color1 = "#007FA6"
Color2 = "#007FA6" 
Color.l = "black" 
Color.l2 = "#878787"
Color.up = "#CB326D"
Color.down = "#00AFBB" 
font = "Calibri"
footer = NULL
    
  
  ## Flextable
  # Set info
  ft <- flextable(Data, col_keys = colnames(Data))
  # Set header
  ft <- set_header_df(ft, mapping = header, key = "col_keys" )
  # Format number
  ft <- colformat_num(ft, big.mark = ".", decimal.mark = "," )
  # Set footer
  if(! is.null(footer)){
    ft <- add_footer(ft, values = footer) %>% 
      #merge_h(part = "footer")  %>% 
      bold(j=1, part = "footer") %>%
      hline(border = fp_border(width = 2, color = Color1), part = "footer")
  }
  
  # Merge headers
  ft <- merge_h(ft, part = "header") %>% 
    merge_v( part = "header")
  
  # Merge contrastes
  #ft <- merge_v(ft, j = c(1,2), part = "body")
  
  # Horizontal lines 
  ft <- hline(ft,border = fp_border(width = 2, color = Color1), part = "header")
  
  # Outer bottom
  ft <- hline_bottom(ft, border = fp_border(width = 2, color = Color1))
  
  # Alineamiento
  ft <- flextable::align(ft, align = "center", part = "header")
  ft <- flextable::align(ft, align = "center", part = "footer")
  ft <- flextable::align(ft, align = "center", part = "body")
  
  # Primera columna alineada
  ft <- flextable::align(ft, j=1, align = "left", part = "all")
  
  
  # Header design 
  x = Data[, 1]
    colors = case_when(
    x == "Up" ~ Color.up,
    x == "Down" ~ Color.down,
    TRUE ~ Color.l2
  )
  
  ft <- fontsize(ft, i = 1:2, size = 12, part = "header") %>%
    color(i=1, color = Color2, part = "header")  %>%
    color(i=c(2), color = Color.l2, part = "header") %>%
    bold( bold = TRUE, part = "header") %>%
    bold( bold = TRUE, j=1, part = "body")  %>%
    # Colores chachis x
    color(j=1, color = colors, part = "body")
  # Rotate names
  #ft <- rotate(ft, i = 2, rotation = "btlr", part = "header", align = "bottom")
  
  # Font
  #
  
  # Color matrix
   ft<- fontsize(ft,  size = 12, part = "all") %>%
     font(fontname = font, part = "all")
ft <-ft %>% flextable::width(width = 0.3)

  # Footer
  if(! is.null(footer)){
    ft <- add_footer(ft, values = footer) %>% 
      merge_h(part = "footer")  %>% 
      bold(j=1, part = "footer") %>%
      hline(border = fp_border(width = 2, color = Color1), part = "footer")
      ft <- flextable::align(ft, align = "right", part = "footer")
  ft <- flextable::align(ft, j=1, align = "left", part = "footer")
  }

# We paint the values grey -
  colormatrix <- ifelse((as.matrix(Data)) %in% c(0,"0", "-"), Color.l2, Color.l )
ft <- color(ft, color = colormatrix, part = "body")  %>%
    color(j=1, color = "black", part = "body")
ft
```


## Obesity {-}

The following results were obtained in the analysis on the basis of the Biological processes of GO:

```{r echo=F, eval=preloadData, message=FALSE, warning=FALSE}
## Obesity
# Get data
Niveles = c("IO" = "Ob - Np",
            "IOM"="Ob.M - Np.M",
            "IOF" = "Ob.F - Np.F",
            "SDIO" = "(Ob.M - Np.M) - (Ob.F-  Np.F)")
Results = get(load(glue("{DirOut}/Obesity_gseBP.RData")))
names(Results) = names(Niveles)
Resultsob = sapply(Results, function(x)
  data.frame(x), simplify = FALSE)

Rterms = get(load(glue("{DirOut}/Ob_reducedTerms.RData")))
RtermsDM = Rterms[c("IOM", "IOF")]
```  

```{r echo=T, eval=loadData, message=FALSE, warning=FALSE}
## Obesity
# Get data
Niveles = c("IO" = "Ob - Np",
            "IOM"="Ob.M - Np.M",
            "IOF" = "Ob.F - Np.F",
            "SDIO" = "(Ob.M - Np.M) - (Ob.F-  Np.F)")
Results = get(load(glue("{DirOut}/Obesity_gseBP.RData")))
names(Results) = names(Niveles)
Resultsob = sapply(Results, function(x)
  data.frame(x), simplify = FALSE)

# Count
simMatrix_IOM <- calculateSimMatrix(Resultsob$IOM$ID,
                                    orgdb="org.Hs.eg.db",
                                    ont="BP",
                                    method="Rel")
simMatrix_IOF <- calculateSimMatrix(Resultsob$IOF$ID,
                                    orgdb="org.Hs.eg.db",
                                    ont="BP",
                                    method="Rel")
# We reduce the terms in  males
scores_IOM <- setNames(-log10(Resultsob$IOM$p.adjust), Resultsob$IOM$ID)
reducedTerms_IOM <- reduceSimMatrix(simMatrix_IOM,
                                    scores_IOM,
                                    threshold=0.9,
                                    orgdb="org.Hs.eg.db")
TB_IOM = sort(table(reducedTerms_IOM$parentTerm), decreasing = TRUE)
#table(cumsum(TB_IOM) > sum(TB_IOM)*0.75) 

# Obtaining the similarity matrices
scores_IOF <- setNames(-log10(Resultsob$IOF$p.adjust), Resultsob$IOF$ID)
reducedTerms_IOF <- reduceSimMatrix(simMatrix_IOF,
                                    scores_IOF,
                                    threshold=0.9,
                                    orgdb="org.Hs.eg.db")
sort(table(reducedTerms_IOF$parentTerm), decreasing = TRUE)
TB_IOF = sort(table(reducedTerms_IOF$parentTerm), decreasing = TRUE)
#table(cumsum(TB_IOF) > sum(TB_IOF)*0.75)

#Final data
RtermsOB = list(IOM = reducedTerms_IOM, IOF = reducedTerms_IOF) 
save(RtermsOB, file=glue("{DirOut}/Ob_reducedTerms.RData"))
```  

### Dotplot 

Dotplot of the most over- and under-expressed terms in the IOM, IOF and SDIO contrasts:  
  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, fig.align='center', fig.width= 10, fig.height= 8}
# Data
Results = Resultsob
Niv = c("IO", "IOM", "IOF", "SDIO" )
names(Results) = Niv

# We took those with the most over- and under-expression
# in females, males and SDIO
# IOM
DfM = data.frame(Results[[Niv[[2]]]])
DfM = DfM[order(DfM$NES, decreasing = T), ]
# We take the 5 most overexpressed and underexpressed ones
nup = ifelse(length(DfM[DfM$NES > 0,]) >5, 5, length(DfM[DfM$NES > 0,]))
ndown = ifelse(length(DfM[DfM$NES < 0,]) >5, 5, length(DfM[DfM$NES < 0,]))
M_dots <- DfM[c(1:nup, (nrow(DfM)-(ndown-1)):nrow(DfM)), ]$ID

# IOF
DfF = data.frame(Results[[Niv[[3]]]])
DfF = DfF[order(DfF$NES, decreasing = T), ]
# We take the 5 most overexpressed and underexpressed ones
nup = ifelse(length(DfF[DfF$NES > 0,]) >5, 5, length(DfF[DfF$NES > 0,]))
ndown = ifelse(length(DfF[DfF$NES < 0,]) >5, 5, length(DfF[DfF$NES < 0,]))
F_dots <- DfF[c(1:nup, (nrow(DfF)-(ndown-1)):nrow(DfF)), ]$ID

# SDIO
DfSD = data.frame(Results[[Niv[[4]]]])
DfSD = DfSD[order(DfSD$NES, decreasing = T), ]
# We take the 5 most overexpressed and underexpressed ones
nup = ifelse(length(DfSD[DfSD$NES > 0,]) >5, 5, length(DfSD[DfSD$NES > 0,]))
ndown = ifelse(length(DfSD[DfSD$NES < 0,]) >5, 5, length(DfSD[DfSD$NES < 0,]))
SD_dots <- DfSD[c(1:nup, (nrow(DfSD)-(ndown-1)):nrow(DfSD)), ]$ID

# We aggregate the data
Dots <- unique(c(M_dots, F_dots, SD_dots)) 

DataM <- DfM[which(DfM$ID %in% Dots),]
DataM$Cluster <- as.factor(Niv[[2]])

DataF <- DfF[which(DfF$ID %in% Dots),]
DataF$Cluster <- as.factor(Niv[[3]])

DataSD <- DfSD[which(DfSD$ID %in% Dots),]
DataSD$Cluster <- as.factor(Niv[[4]])

Data = rbind(DataM, DataF, DataSD)
Data$Description = str_wrap(str_to_sentence(Data$Description), width = 40) 
Data <- Data %>%
  arrange(p.adjust, desc(abs(NES)))
Data$Direction = ifelse(Data$NES < 0, "logOR<0", "logOR>0") 
Data$Description = factor((Data$Description),
                          levels=unique((Data$Description))) 

# DotPlot
DotP = ggplot(Data, aes( reorder(Description, rev(p.adjust)),
                        Direction)) +
  geom_point(shape = 21,
             aes(fill = NES, 
                 size = setSize,
                 color = NES, 
                 alpha = p.adjust)) +  
  facet_grid(~Cluster) + coord_flip() + xlab("GO: Biological processes") + 
  ylab("") +
  scale_fill_gradient(low="#90C433", high="#2494b6") +
  scale_alpha(range = c(0.8, 0.2)) + 
  geom_point(shape = 21, stroke = 0.5,
             aes(color = NES, 
                 size = setSize),
             fill = "transparent") +
  scale_color_gradient(low="#90C433", high="#2494b6", guide = "none") +
  guides(alpha=guide_legend(override.aes=list(fill=hcl(c(15,195),100,0,
                                                       alpha=c(0.01, 0.9)),
                                              size = 4,
                                              colour="#878787"),
                            title = "FDR",
                            order = 2),
         size = guide_legend(title = "Set size",
                            order = 3),
         fill = guide_colorbar(title = "logOR",
                               order = 1)) + 
    theme(plot.background = element_rect(fill="transparent"),
          panel.background = element_rect(fill = "grey95"),
          axis.line = element_line(colour = "grey"))
DotP 
```

### REVIGO 

The results for the grouping of the terms based on distance are shown below
based on distance:

#### IOM 

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, fig.align='center', fig.width= 10, fig.height= 8}
# Representation of reduced terms
RX_revigo(RtermsOB$IOM) 
```

#### IOF 

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, fig.align='center', fig.width= 10, fig.height= 8}
# Representation of reduced terms 
RX_revigo(RtermsOB$IOF)
```

## Diabetes {-}  

The following results were obtained in the analysis on the basis of the Biological processes of GO:

```{r echo=F, eval=preloadData, message=FALSE, warning=FALSE}
## Diabetes
# Get data
Niveles = c("ID" = "Ob - Np",
            "IDM"="Ob.M - Np.M",
            "IDF" = "Ob.F - Np.F",
            "SDID" = "(Ob.M - Np.M) - (Ob.F-  Np.F)")
Results = get(load(glue("{DirOut}/FA_gseBP.RData")))
names(Results) = names(Niveles)
ResultsDM = sapply(Results, function(x)
  data.frame(x), simplify = FALSE)

Rterms = get(load(glue("{DirOut}/T2D_reducedTerms.RData")))
RtermsDM = Rterms[c("IDM", "IDF")]
```  


```{r echo=T, eval=loadData, message=FALSE, warning=FALSE}
## Diabetes
# Get data
Niveles = c("ID" = "Ob - Np",
            "IDM"="Ob.M - Np.M",
            "IDF" = "Ob.F - Np.F",
            "SDID" = "(Ob.M - Np.M) - (Ob.F-  Np.F)")
Results = get(load(glue("{DirOut}/FA_gseBP.RData")))
names(Results) = names(Niveles)
ResultsDM = sapply(Results, function(x)
  data.frame(x), simplify = FALSE)

# Obtaining the similarity matrices
simMatrix_IDM <- calculateSimMatrix(ResultsDM$IDM$ID,
                                    orgdb="org.Hs.eg.db",
                                    ont="BP",
                                    method="Rel")
simMatrix_IDF <- calculateSimMatrix(ResultsDM$IDF$ID,
                                    orgdb="org.Hs.eg.db",
                                    ont="BP",
                                    method="Rel")

# We reduce the terms in  males
scores_IDM <- setNames(-log10(ResultsDM$IDM$p.adjust), ResultsDM$IDM$ID)
reducedTerms_IDM <- reduceSimMatrix(simMatrix_IDM,
                                    scores_IDM,
                                    threshold=0.75,
                                    orgdb="org.Hs.eg.db")
TB_IDM = sort(table(reducedTerms_IDM$parentTerm), decreasing = TRUE)
#table(cumsum(TB_IDM) > sum(TB_IDM)*0.75)

# We reduce the terms in  females
scores_IDF <- setNames(-log10(ResultsDM$IDF$p.adjust), ResultsDM$IDF$ID)
reducedTerms_IDF <- reduceSimMatrix(simMatrix_IDF,
                                    scores_IDF,
                                    threshold=0.75,
                                    orgdb="org.Hs.eg.db")
TB_IDF = sort(table(reducedTerms_IDF$parentTerm), decreasing = TRUE)
#table(cumsum(TB_IDF) > sum(TB_IDF)*0.75)

#Final data
RtermsDM = list(IDM = reducedTerms_IDM, IDF = reducedTerms_IDF) 
save(RtermsDM, file=glue("{DirOut}/T2D_reducedTerms.RData"))
```  

### Dotplot 

Dotplot of the most over- and under-expressed terms in the contrasts 
IDM, IDF and SDID:  
  

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, fig.align='center', fig.width= 10, fig.height= 8}
# Data
Results = ResultsDM
Niv = c("ID", "IDM", "IDF", "SDID" )
names(Results) = Niv

# We took those with the most over- and under-expression
# in females, males and SDID
# IDM
DfM = data.frame(Results[[Niv[[2]]]])
DfM = DfM[order(DfM$NES, decreasing = T), ]
nup = ifelse(length(DfM[DfM$NES > 0,]) >5, 5, length(DfM[DfM$NES > 0,]))
ndown = ifelse(length(DfM[DfM$NES < 0,]) >5, 5, length(DfM[DfM$NES < 0,]))
M_dots <- DfM[c(1:nup, (nrow(DfM)-(ndown-1)):nrow(DfM)), ]$ID

# IDF
DfF = data.frame(Results[[Niv[[3]]]])
DfF = DfF[order(DfF$NES, decreasing = T), ]
# We take the 5 above and below
nup = ifelse(length(DfF[DfF$NES > 0,]) >5, 5, length(DfF[DfF$NES > 0,]))
ndown = ifelse(length(DfF[DfF$NES < 0,]) >5, 5, length(DfF[DfF$NES < 0,]))
F_dots <- DfF[c(1:nup, (nrow(DfF)-(ndown-1)):nrow(DfF)), ]$ID

# SDID
DfSD = data.frame(Results[[Niv[[4]]]])
DfSD = DfSD[order(DfSD$NES, decreasing = T), ]
# We take the 5 above and below
nup = ifelse(length(DfSD[DfSD$NES > 0,]) >5, 5, length(DfSD[DfSD$NES > 0,]))
ndown = ifelse(length(DfSD[DfSD$NES < 0,]) >5, 5, length(DfSD[DfSD$NES < 0,]))
SD_dots <- DfSD[c(1:nup, (nrow(DfSD)-(ndown-1)):nrow(DfSD)), ]$ID

# We aggregate the data
Dots <- unique(c(M_dots, F_dots, SD_dots)) 

DataM <- DfM[which(DfM$ID %in% Dots),]
DataM$Cluster <- as.factor(Niv[[2]])

DataF <- DfF[which(DfF$ID %in% Dots),]
DataF$Cluster <- as.factor(Niv[[3]])

DataSD <- DfSD[which(DfSD$ID %in% Dots),]
DataSD$Cluster <- as.factor(Niv[[4]])

Data = rbind(DataM, DataF, DataSD)
Data$Description = str_wrap(capitalize(Data$Description), width = 50) 
Data <- Data %>%
  arrange(p.adjust, desc(abs(NES)))
Data$Direction = ifelse(Data$NES < 0, "logOR<0", "logOR>0") 
Data$Description = factor((Data$Description), levels=unique((Data$Description))) 

# DotPlot
DotP = ggplot(Data, aes(reorder(Description, rev(p.adjust)), 
                        Direction)) +
  geom_point(shape = 21,
             aes(fill = NES, 
                 size = setSize,
                 color = NES, 
                 alpha = p.adjust)) +  
  facet_grid(~Cluster) + coord_flip() + xlab("GO: Biological processes") + ylab("") +
  scale_fill_gradient(low="#2494b6", high="#90C433") +
  scale_alpha(range = c(0.8, 0.2)) + 
  geom_point(shape = 21, stroke = 0.5,
             aes(color = NES, 
                 size = setSize),
             fill = "transparent") +
  scale_color_gradient(low="#2494b5", high="#90C432", guide = "none") +
  guides(alpha=guide_legend(override.aes=list(fill=hcl(c(15,195),100,0,
                                                           alpha=c(0.01, 0.9)),
                                              size = 4,
                                              colour="#878787"),
                            title = "FDR",
                            order = 2),
         size = guide_legend(title = "Set size",
                            order = 3),
         fill = guide_colorbar(title = "logOR",
                               order = 1)) + 
    theme(plot.background = element_rect(fill="transparent"),
          panel.background = element_rect(fill = "grey95"),
          axis.line = element_line(colour = "grey"))
DotP 
```

### REVIGO 

The results for the clustering of terms based on distance are shown below:

#### IDM 

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, fig.align='center', fig.width= 10, fig.height= 10}
# Representation of reduced terms
RX_revigo(RtermsDM$IDM)  
```

#### IDF 

```{r echo=T, eval=TRUE, message=FALSE, warning=FALSE, fig.align='center', fig.width= 10, fig.height= 10}
# Representation of reduced terms 
RX_revigo(RtermsDM$IDF)
```

***      

&nbsp;  

